---
title: 'Generate Test DB'
date: "`r Sys.time()`"
execute:
  message: FALSE
  echo: FALSE
---

```{r}
library(tidyr)
library(dplyr)
library(DBI)
library(here)
library(brio)
library(lubridate)
library(lorem)
library(stringr)

```


```{r Drop and Create Tables}

# Drop and Create Tables
TEST <- dbConnect(RMySQL::MySQL(),
                dbname = "test",
                host = Sys.getenv("DB_HOST"),
                port = 3306,
                user = "admin",
                password = Sys.getenv("DB_PASSWORD"))


sql_script <- readLines("new_db_structure.sql") |> paste(collapse = "\n")

sql_statements <- unlist(strsplit(sql_script, ";")) |> trimws()

# Execute each statement
for (stmt in sql_statements) {
  if (stmt != "") { # Skip empty statements
    tryCatch({
      dbExecute(TEST, stmt)
    }, error = function(e) {
      message("Error executing statement: ", stmt, "\n", e$message)
    })
  }
}


```

```{r Populate Table Functions}

db_insert <- function(table, data) {
  dbWriteTable(TEST, table, data, append = TRUE, row.names = FALSE)
}

# Helper to generate random dates near today
generate_date <- function(offset_days = 0, range_days = 30, past_only = TRUE) {
  today <- Sys.Date()
  if(past_only) {
    as.character(today + offset_days + sample(-range_days:0, 1))
  } else {
    as.character(today + offset_days + sample(-range_days:range_days, 1))
  }
  
}


# Helper to generate unique IDs
generate_id <- function(prefix, n = 6) {
  paste0(prefix, sprintf("%0*d", n, sample(1:10^n, 1)))
}

# Helper to generate address
generate_address <- function() {
  paste(sample(1:1000, 1), 
        sample(c('Spuce', 'Elm', 'Pine',
                 'Douglas', 'Maple', 'Juniper',
                 'Cedar', 'Birch', 'Willow',
                 'Oak', 'Hickory', 'Walnut',
                 'Cherry', 'Pecan', 'Chestnut',
                 'Sycamore', 'Poplar', 'Aspen',
                 'Beech', 'Ash', 'Alder',
                 'Locust', 'Cottonwood', 'Magnolia',
                 'Palm', 'Redwood', 'Sequoia'), 1),
        sample(c('St', 'Ave', 'Blvd', 'Ct', 'Dr', 'Ln', 'Rd', 'Way'), 1)
        )
}


```

## Populate `company`
```{r company}
company <- data.frame(
  company_id = 1:3,
  company_name = c("Admin", "A", "B")
)

db_insert("company", company)
```

## Populate `firefighter` and `chain_of_command`
```{r firefighter}

firefighter <- data.frame(
  firefighter_id = 1:11,
  full_name = c(
    "Bill Preston", "Pete Mitchell", "Napolean Dynamite", "Lane Meyer", "Steve Rogers",
    "Samwise Gamgee", "Scott Pilgram", "Angela Bennett", "Thomas Anderson", "George Bailey", "Marty McFly"
  ),
  start_date = c(
    "1989-02-17", "1986-05-16", "2004-06-11", "1985-10-11", "2011-07-22",
    "2001-12-19", "2010-08-13", "1985-07-28", "1999-03-31", "1946-12-20", "1985-07-03"
  ),
  trainer = sample(0:1, 11, replace = TRUE, prob = c(0.2, 0.8)),
  officer = c(0, 0, 1, 0, 1, 0, 1, 0, 0, 0, 1),
  active_status = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 1),
  company_id = c(2, 3, 1, 2, 2, 3, 1, 3, 2, 2, 3),
  firefighter_role = c("Lieutenant", "Firefighter", "Chief", "Probationary", "Captain", "Lieutenant", "Assistant Chief", "Firefighter", "Firefighter", "Firefighter", "Captain"),
  supervisor = c(5, 11, NA, 5, 3, 11, 3, 11, 5, NA, 3)
)

chain_of_command <- firefighter |> 
  mutate(
    chain_of_command_id = 1:11,
    firefighter_id = firefighter_id,
    supervisor_id = supervisor,
    .keep = "none"
  ) |> 
  relocate("chain_of_command_id", .before = "firefighter_id")

firefighter <- firefighter |> 
  select(-supervisor)

db_insert("firefighter", firefighter)
db_insert("chain_of_command", chain_of_command)
```


## Populate `firefighter_contact`
```{r firefighter_contact}
firefighter_contact <- data.frame(
  firefighter_id = 1:11,
  street_address_1 = sapply(1:11, function(x) generate_address()),
  street_address_2 = NA,
  city = "Crabapple Cove",
  state = "ME",
  zip_code = "04578",
  phone_number = sapply(1:11, function(x) paste0("(", sample(100:999, 1), ")-555-", sample(1000:9999, 1))),
  email_address = sapply(1:11, function(x) paste0(tolower(str_replace(firefighter$full_name[x], " ", "")), "@firedept.com"))
  )

firefighter_contact[3,3] <- "#65"

db_insert("firefighter_contact", firefighter_contact)

```

## Populate `firefighter_status_history`
```{r firefighter_status_history}
firefighter_status_history <- data.frame(
  firefighter_id = sample(1:9, 2, replace = FALSE),
  inactive_start_date = sapply(1:2, function(x) generate_date(-730, 180)),
  inactive_end_date = sapply(1:2, function(x) generate_date(-365, 180))
)


firefighter_status_history[3,] <- c(10, "2010-01-01", NA)


db_insert("firefighter_status_history", firefighter_status_history)

```

## Populate `certification_type`
```{r certification_type}

certification_type <- data.frame(
  certification_type_id = 1:6,
  certification_name = c("Firefighter I", "Firefighter II", "EMT", "AEMT", "CPR", "Wildland FFT2"),
  lead_time = c(6, 6, 6, 6, 6, 6),
  lead_time_unit = "month",
  renew_time = c(3, 3, 2, 2, 2, 1),
  renew_time_unit = "year",
  certification_expire = c(rep(NA, 5), "2025-01-01")
)

db_insert("certification_type", certification_type)


```

## Populate `certification`
```{r certification}
certification <- data.frame(
  certification_id = 1:25,
  type_id = sample(1:6, 25, replace = TRUE),
  firefighter_id = sample(1:11, 25, replace = TRUE),
  expiration_date = sapply(1:25, function(x) generate_date(180, 365, FALSE)),
  snooze_expires = sapply(1:25, function(x) generate_date(180, 365, FALSE))
)

db_insert("certification", certification)


```

## Populate `setting`
```{r setting}

setting <- read.csv("settings.csv")

db_insert("setting", setting)

```

## Populate `user_log`
```{r user_log}

user_log <- data.frame(
  user_log_id = 1:10,
  log_in = sapply(1:10, function(x) 
    ymd_hms(
      paste(
        generate_date(-7, 0),
        paste(sample(0:23, 1, TRUE)),
        paste(sample(0:59, 2, TRUE), collapse = ":")
        )
      )
    )
  ) |> 
  mutate(log_out = log_in + sample(60:360, 10, replace = TRUE),
         log_in = as.POSIXct(log_in) |> format(usets = FALSE),
         log_out = as.POSIXct(log_out) |> format(usets = FALSE))

db_insert("user_log", user_log)

```

## Populate `apparatus`
```{r}
apparatus <- data.frame(
  apparatus_id = 1:6,
  apparatus_name = c("Engine 1", "Ladder 2", "Rescue 3", "Ambulance 1", "Battalion 1", "POV")
)
db_insert("apparatus", apparatus)
```

## Populate `equipment_type`
```{r}
equipment_type <- data.frame(
  equipment_type_id = 1:4,
  equipment_type = c("SCBA Battery", "Apparatus", "Turnout Gear", "DO NOT USE"),
  check_lead_time = c(7, 7, 1, 1),
  check_lead_time_unit = c("day", "day", "year", "year"),
  check_time = c(1, 1, 1, 1),
  check_time_unit = c("month", "month", "year", "year"),
  expire_lead_time = c(6, NA, 1, 1),
  expire_lead_time_unit = c("month", NA, "year", "year"),
  expire_time = c(3, NA, 10, 5),
  expire_time_unit = c("year", NA, "year", "year"),
  equipment_type_expire = c(rep(NA, 3), "2025-01-01")
)

db_insert("equipment_type", equipment_type)

```

## Populate `equipment`
```{r}

equipment <- data.frame(
  equipment_id = 1:8,
  equipment_name = c("Engine 1 SCBA Batteries", "Engine 2 SCBA Batteries", "Ladder 1 SCBA Batteries", 
                     "Engine 1 Starts", "Ladder 2 Starts",
                     "Turnouts", "Turnouts", "Turnouts"),
  equipment_type_id = c(1,1,1,2,2,3,3,3),
  firefighter_id = c(NA,NA,NA,NA,NA,1,3,10),
  next_check_date = c(
    Sys.Date() - 3,
    Sys.Date() + 5,
    Sys.Date() + 17,
    Sys.Date() - 3,
    Sys.Date() + 8,
    Sys.Date() + 180,
    Sys.Date() + 180,
    Sys.Date() + 180
  ),
  expiration_date = c(
    Sys.Date() - 100,
    Sys.Date() + 365,
    Sys.Date() + 800,
    Sys.Date() + 300,
    as.Date(NA),
    Sys.Date() + 180,
    Sys.Date() + 180,
    Sys.Date() + 720
  ),
  snooze_expires = c(
    as.Date(NA),
    as.Date(NA),
    as.Date(NA),
    as.Date(NA),
    as.Date(NA),
    Sys.Date() + 180,
    Sys.Date() + 180,
    Sys.Date() + 720
  )
)

db_insert("equipment", equipment)

```

## Populate `equipment_log`
```{r}

equipment_log <- data.frame(
  equipment_log_id = 1:10,
  checked_by_id = sample(1:11, 10, replace = TRUE),
  equipment_id = sample(1:10, 10, replace = TRUE),
  check_date = sapply(1:10, function(x) 
    ymd_hms(
      paste(
        generate_date(-7, 0),
        paste(sample(0:23, 1, TRUE)),
        paste(sample(0:59, 2, TRUE), collapse = ":")
        )
      )
    )
  ) |> 
  mutate(
    check_date = as.POSIXct(check_date) |> format(usets = FALSE)
  )

db_insert("equipment_log", equipment_log)

```

## Populate `training`
```{r}
training <- data.frame(
  training_id = 1:50,
  training_type = sample(c('Fire', 'EMS', 'Wildland', 'Other'), 50, replace = TRUE, prob = c(0.4, 0.4, 0.15, 0.05)),
  topic = sample(lorem::ipsum_words(n = 6) |> stringr::str_split(" ") |> unlist(), 50, replace = TRUE),
  training_description = sapply(1:05, function(x) lorem::ipsum_words(n = 6)),
  start_time = sapply(1:50, function(x) 
    ymd_hms(
      paste(
        generate_date(0, 365),
        paste(sample(c(0:3), 1, TRUE)),
        paste(sample(c(0,30), 1, TRUE),
              "00",
              collapse = ":")
        )
      )
    ),
  trainer = sample(1:11, 50, replace = TRUE),
  training_delete = NA
) |> 
  mutate(end_time = start_time + 60 * sample(c(90, 120, 150), 50, replace = TRUE),
         .after = "start_time",
         start_time = as.POSIXct(start_time) |> format(usets = FALSE),
         end_time = as.POSIXct(end_time) |> format(usets = FALSE))


db_insert("training", training)
```

## Populate `attendance`
```{r}
loading_df <- data.frame()

for (x in training$training_id) {
  n <- sample(1:11, 1)
  firefighter_ids <- sample(1:11, n, replace = FALSE)
  check_in_times <- sapply(1:n, function (i) 
    training[x, "start_time"] |> as.POSIXct() + sample(-5:45, 1) * 60
  )
  
  
  attendance <- data.frame(
    firefighter_id = firefighter_ids,
    training_id = rep(x, n),
    check_in = check_in_times |> as.POSIXct() |>  format(usets = FALSE),
    check_out = (check_in_times |> as.POSIXct() + sample(60:120, n, replace = TRUE) * 60) |>  format(usets = FALSE),
    auto_checkout = sample(0:1, n, replace = TRUE, prob = c(0.9, 0.1)),
    credit = sample(0:1, n, replace = TRUE, prob = c(0.02, 0.98)),
    excused = sample(0:1, n, replace = TRUE, prob = c(0.95, 0.05))
  )
  
  attendance$check_out[attendance$excused == 1] <- NA
  attendance$check_in[attendance$excused == 1] <- NA
  attendance$auto_checkout[attendance$excused == 1] <- NA
  attendance$credit[attendance$excused == 1] <- NA
  
  
  loading_df <- bind_rows(loading_df, attendance)
  
}

loading_df$attendance_id <- 1:nrow(loading_df)

db_insert("attendance", loading_df)

```


## Populate `incident` and `incident_xref`
```{r}
dispatch_codes <- c(
  "Abdominal Pain/Problems",
  "Allergies / Envenomations",
  "Animal Bites / Attacks",
  "Assault / Sexual Assault / Stun Gun",
  "Back Pain",
  "Breathing Problems",
  "Burns / Explosions",
  "Carbon Monoxide / Inhalation / HAZMAT / CBRN",
  "Cardiac or Respiratory Arrest / Death",
  "Chest Pain",
  "Choking",
  "Convulsions / Seizures",
  "Diabetic Problems",
  "Drowning / Diving / SCUBA Accident",
  "Electrocution / Lightning",
  "Eye Problems / Injuries",
  "Falls",
  "Headache",
  "Heart Problems / A.I.C.D.",
  "Heat / Cold Exposure",
  "Hemorrhage / Lacerations",
  "Inaccessible Incident / Entrapments",
  "Overdose / Poisoning (Ingestion)",
  "Pregnancy / Childbirth / Miscarriage",
  "Psychiatric / Suicide Attempt",
  "Sick Person",
  "Stab / Gunshot / Penetrating Trauma",
  "Stroke (CVA) / Transient Ischemic Attack (TIA)",
  "Traffic / Transportation Incidents",
  "Traumatic Injuries",
  "Unconscious / Fainting (Near)",
  "Unknown Problem",
  "Inter-Facility Transfer / Palliative Care",
  "Automatic Crash Notification (A.C.N.)",
  "Pandemic / Epidemic / Outbreak (Surveillance or Triage)",
  "CO Alarm",
  "Fire Alarm",
  "Garage Fire",
  "Grass Fire",
  "HAZMAT",
  "Illegal Burn",
  "Motor Vehicle Accident (Injuries)",
  "Motor Vehicle Accident (No Injuries)",
  "Smoke Investigation",
  "Structure Fire",
  "Vehicle Fire"
)

incident <- data.frame(
  incident_id = sapply(1:50, function(x) generate_id("INC")),
  dispatch_time = sapply(1:50, function(x) 
    ymd_hms(
      paste(
        generate_date(0, 60),
        paste(sample(0:23, 1, TRUE)),
        paste(sample(0:59, 2, TRUE), collapse = ":")
        )
      )
    ),
  address = sapply(1:50, function(x) generate_address()),
  dispatch_reason = sample(dispatch_codes, 50, TRUE),
  ems_units = sample(0:1, 50, replace = TRUE),
  fire_units = sample(0:1, 50, replace = TRUE),
  wildland_units = sample(0:1, 50, replace = TRUE),
  area = sample(c('Municipality', 'Primary Area', 'Mutual Aid', 'Outside Aid'), 50, replace = TRUE),
  canceled = sample(0:1, 5, replace = TRUE, prob = c(0.9, 0.1)),
  dropped = sample(0:1, 5, replace = TRUE, prob = c(0.95, 0.05)),
  notes = NA
  # finalized = sample(0:1, 5, replace = TRUE)
) |> 
  mutate(end_time = dispatch_time + sample(300:36000, 50, replace = TRUE),
         .after = "dispatch_time") |> 
  mutate(finalized = if_else(end_time > Sys.time() - 60*60*24*7, 0, 1),
         dispatch_time = as.POSIXct(dispatch_time) |> format(usets = FALSE),
         end_time = as.POSIXct(end_time)|> format(usets = FALSE)) |> 
  arrange(desc(end_time))

top_incident <- incident|> 
  slice(1)

additional_incident <- incident|> 
  slice(2) |> 
  mutate(address = NA,
         dispatch_reason = NA,
         ems_units = NA,
         fire_units = NA,
         wildland_units = NA,
         area = NA,
         canceled = NA,
         dropped = NA,
         incident_id = paste0(top_incident$incident_id, "-additional-50")
         )

incident <- incident|> 
  slice(-2) |> 
  bind_rows(additional_incident)


incident_xref <- data.frame(
  id = 1,
  primary_incident = top_incident$incident_id,
  additional_incident = paste0(top_incident$incident_id, "-additional-50")
)

db_insert("incident", incident)
db_insert("incident_xref", incident_xref)
```

## Populate `firefighter_incident`, `apparatus_incident`, and `firefighter_apparatus`
```{r}
loading_firefighter_incident <- data.frame()
loading_apparatus_incident <- data.frame()
loading_firefighter_apparatus <- data.frame()

for (x in incident$incident_id) {
  n <- sample(1:11, 1)
  m <- sample(1:5, 1)
  
  m <- min(m, n)
  
  firefighter_ids <- sample(1:11, n, replace = FALSE)
  apparatus_ids <- sample(1:5, m, replace = FALSE)
  
  firefighter_incident <- data.frame(
    incident_id = rep(x, n),
    firefighter_id = firefighter_ids,
    time_adjustment = sample(c(-60, -30, 0, 30, 60), 
                             n, replace = TRUE, 
                             prob = c(0.025, 0.05, 0.85, 0.05, 0.025))
  )
  
  apparatus_incident <- data.frame(
    incident_id = rep(x, m),
    apparatus_id = apparatus_ids,
    time_adjustment = sample(c(-60, -30, 0, 30, 60), 
                             m, replace = TRUE, 
                             prob = c(0.025, 0.05, 0.85, 0.05, 0.025))
  )
  
  firefighter_apparatus <- data.frame(
    incident_id = rep(x, n),
    firefighter_id = sample(firefighter_ids, n, replace = FALSE),
    apparatus_id = sample(apparatus_ids, n, replace = TRUE)
  )
  
  loading_firefighter_incident <- bind_rows(loading_firefighter_incident, firefighter_incident)
  loading_apparatus_incident <- bind_rows(loading_apparatus_incident, apparatus_incident)
  loading_firefighter_apparatus <- bind_rows(loading_firefighter_apparatus, firefighter_apparatus)

}

loading_firefighter_incident$firefighter_incident_id <- 1:nrow(loading_firefighter_incident)
loading_apparatus_incident$apparatus_incident_id <- 1:nrow(loading_apparatus_incident)
loading_firefighter_apparatus$firefighter_apparatus_id <- 1:nrow(loading_firefighter_apparatus)

db_insert("firefighter_incident", loading_firefighter_incident)
db_insert("apparatus_incident", loading_apparatus_incident)
db_insert("firefighter_apparatus", loading_firefighter_apparatus)

```



## Disconnect from Database
```{r}
dbDisconnect(TEST)
```


